---
title: "FeedingSocialTolerance"
author: "Alba Motes Rodrigo"
date: "2026-03-05"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(igraph)
```


```{r}
#load dates of carpet experiments
carpet_dates<-read.csv("/Users/alba/Library/Mobile Documents/com~apple~CloudDocs/Postdoc_UNIL/Vervet_projects/FeedingSocialityProject/carpet_dates.csv")
#load scans
scans<-read.csv("/Users/alba/Library/Mobile Documents/com~apple~CloudDocs/Postdoc_UNIL/Vervet_projects/IVP_data/scans/scanswithid_ivp_filtered.csv")
```

```{r}
scans <- scans %>%
  mutate(
    month_day   = as.numeric(format(as.Date(date), "%m%d")),
    season      = ifelse(month_day >= 515 & month_day <= 1115, "wet", "dry"),
    season_id   = case_when(
      month_day >= 515 & month_day <= 1115 ~ paste("wet", year, sep = "_"),
      month_day > 1115 ~ paste("dry", year, sep = "_"),
      month_day < 515  ~ paste("dry", year - 1, sep = "_")
    )
  ) %>%
  dplyr::select(-month_day)
```

```{r}
#calculate edgelist with custom function
source("/Users/alba/Library/Mobile Documents/com~apple~CloudDocs/Postdoc_UNIL/Vervet_projects/FeedingSocialityProject/Anlyses/functions/create_edglistarm1m_fromscans.R")
edgelist_scans <- create_scan_edgelist(scans)
#Remove cases where one dyad member did not appear in any scan of that group that year
edgelist_scans <- edgelist_scans %>% filter(!is.na(scans_individual_a) & !is.na(scans_individual_b))
```

#Function to calcualte individual network metrics from the edgelist of 1m and arm length scans
individual_strength = sum(co_occurrence_scans per dyad) / scans_sampled
scans_sampled = number of scans in which that individual appears
co_occurrence = number of scans where both individuals appear
```{r}
calculate_network_metrics <- function(data, grp,  ssn) {
  
  subset <- data %>% filter(group == grp, season_id == ssn)
  
  if (nrow(subset) == 0) return(NULL)
  
  # Build weighted graph
  g <- graph_from_data_frame(
    subset %>% dplyr::select(individual_a, individual_b, weight = association_rate),
    directed = FALSE
  )
  
  # Reshape to long format so each individual appears once per dyad
  individual_dyads <- bind_rows(
    subset %>% dplyr::select(individual = individual_a, partner = individual_b,
                      co_occurrence_scans, association_rate, scans_sampled = scans_individual_a),
    subset %>% dplyr::select(individual = individual_b, partner = individual_a,
                      co_occurrence_scans, association_rate, scans_sampled = scans_individual_b)
  )
  
  # Individual-level stats
  individual_stats <- individual_dyads %>%
    group_by(individual) %>%
    summarise(
      scans_sampled = unique(scans_sampled),
      degree        = n(),
      strength      = sum(association_rate),
      .groups = "drop"
    )
  
  # Graph-based metrics
  graph_metrics <- data.frame(
    individual  = V(g)$name,
    betweenness = betweenness(g, weights = 1/E(g)$weight, normalized = TRUE),
    eigenvector = eigen_centrality(g, weights = E(g)$weight)$vector
  )
  
  # Combine
  individual_metrics <- individual_stats %>%
    left_join(graph_metrics, by = "individual") %>%
    mutate(group = grp, season_id = ssn)
  
  # Group-level metrics
  group_metrics <- data.frame(
    group            = grp,
    season_id           = ssn,
    n_individuals    = vcount(g),
    n_dyads          = ecount(g),
    total_dyad_scans = sum(subset$total_dyad_scans),
    density          = edge_density(g),
    mean_strength    = mean(strength(g, weights = E(g)$weight)),
    clustering_coef  = transitivity(g, type = "global")
  )
  
  return(list(individual = individual_metrics, group = group_metrics))
}

# Apply across all group-year-season combinations
combinations <- edgelist_scans %>%
  distinct(group, season_id)

all_individual <- list()
all_group      <- list()

for (i in seq_len(nrow(combinations))) {
  result <- calculate_network_metrics(
    edgelist_scans,
    combinations$group[i],
    combinations$season_id[i]
  )
  if (!is.null(result)) {
    all_individual <- append(all_individual, list(result$individual))
    all_group      <- append(all_group, list(result$group))
  }
}
individual_metrics_df <- bind_rows(all_individual)
group_metrics_df      <- bind_rows(all_group)

```

#subset periods with carpet
```{r}
individual_metrics_filtered <- individual_metrics_df %>%
  filter(group %in% c("Ankhase", "Noha"),
         season_id %in% c("dry_2024", "wet_2025", "dry_2025", "wet_2026"))

group_metrics_filtered <- group_metrics_df %>%
  filter(group %in% c("Ankhase", "Noha"),
         season_id %in% c("dry_2024", "wet_2025", "dry_2025", "wet_2026"))
```

#normalise degree
```{r}
group_size <- data.frame(
  group     = c("Ankhase", "Ankhase", "Ankhase","Ankhase", "Noha",     "Noha",     "Noha", "Noha"),
  season_id = c("dry_2024", "wet_2025", "dry_2025", "wet_2026", "dry_2024", "wet_2025", "dry_2025", "wet_2026"),
  group_size = c(16, 15, 15, 16, 22, 21, 21, 20)
)

individual_metrics_filtered <- individual_metrics_filtered %>%
  left_join(group_size, by = c("group", "season_id")) %>%
  mutate(degree_normalized = degree / (group_size - 1)) 
```

#descriptives number of scans in which individuals were observed
```{r}
tapply(individual_metrics_filtered$scans_sampled, 
       individual_metrics_filtered$season_id, 
       summary)
```
$dry_2024
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    1.0     3.0     5.0     6.3     9.0    16.0 

$wet_2025
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   1.00   19.25   27.50   27.47   34.75   52.00 

$dry_2025
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   1.00   18.00   27.00   27.27   36.00   59.00 

$wet_2026
NULL

#Plot
```{r}
library(ggplot2)
library(tidyr)

# Order seasons chronologically
season_order <- c("dry_2024", "wet_2025", "dry_2025", "wet_2026")

individual_metrics_filtered <- individual_metrics_filtered %>%
  mutate(season_id = factor(season_id, levels = season_order))

group_metrics_filtered <- group_metrics_filtered %>%
  mutate(season_id = factor(season_id, levels = season_order))

# --- Individual metrics: boxplots per season and group ---
individual_long <- individual_metrics_filtered %>%
  pivot_longer(cols = c(strength, degree_normalized, betweenness, eigenvector),
               names_to = "metric", values_to = "value")

ggplot(individual_long, aes(x = season_id, y = value, fill = group)) +
  geom_jitter(aes(color = group), position = position_jitterdodge(jitter.width = 0.2),
              size = 1, alpha = 0.4) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_fill_manual(values = c("Ankhase" = "#E69F00", "Noha" = "#56B4E9")) +
  scale_color_manual(values = c("Ankhase" = "#E69F00", "Noha" = "#56B4E9")) +
  labs(x = "Season", y = "Value", fill = "Group", color = "Group",
       title = "Individual social network metrics per season") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  expand_limits(y = 0)

# --- Group metrics: line plot per season ---
group_long <- group_metrics_filtered %>%
  pivot_longer(cols = c(density, mean_strength, clustering_coef),
               names_to = "metric", values_to = "value")

ggplot(group_long, aes(x = season_id, y = value, color = group, group = group)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_color_manual(values = c("Ankhase" = "#E69F00", "Noha" = "#56B4E9")) +
  labs(x = "Season", y = "Value", color = "Group",
       title = "Group social network metrics per season") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  expand_limits(y = 0)

group_metrics_filtered %>% 
  group_by(group, season_id) %>% 
  filter(n() > 1)
```

```{r}
# How many individuals and dyads in that season?
group_metrics_df %>% 
  filter(group == "Noha", season_id == "dry_2024")

# How many scans total?
scans %>% 
  filter(group == "Noha", season_id == "dry_2024") %>% 
  summarise(n_scans = n_distinct(scan_id), n_individuals = n_distinct(idindividual1))
```
The squashed noha eigenvector centrality for dry_2024 is due to the fact that as there are only 264 scans spread across 64 individuals (median 3 scans), most dyads were never observed in proximity together, giving them an SRI of 0 or near-0 — and eigenvector centrality compounds this by giving low scores to individuals connected only to other low-scoring individuals, pushing almost everyone toward zero.

```{r}
write.csv(group_metrics_filtered, "/Users/alba/Library/Mobile Documents/com~apple~CloudDocs/Postdoc_UNIL/Vervet_projects/FeedingSocialityProject/Results/group_metrics_outexp.csv")
write.csv(individual_metrics_filtered, "/Users/alba/Library/Mobile Documents/com~apple~CloudDocs/Postdoc_UNIL/Vervet_projects/FeedingSocialityProject/Results/individual_metrics_outexp.csv")
```

